---
title: "Assignment 5"
author: "Gracie White"
date: "November 29, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
##Load Packages and Read in the Data 

library(tidyverse)
library(stargazer)
library(kableExtra)
library(RColorBrewer)
library(car)

fac_sal_df <- read_csv("Faculty salary data (2008 - 2009 survey).csv")
grad_enroll_df <- read_csv("Grad Enrollment.csv")
phd_med_sal_df <- read_csv("Median salary for doctoral recipients.csv")
phd_field_df <- read_csv("PhDs by Field 1985 - 2015.csv")
enroll_tidy_df <- read_csv("enrollment tidy.csv")
female_phd_tidy <- read_csv("Female_phds.csv")
female_probtable <- read_csv("phd_proptable.csv")
female_numbers_tidy <- read_csv("female_phd_numbers.csv")
phd_med_sal_tidy_df <- read_csv("PostGrad_Med_Sal_TIDY FORMAT.csv")

```


Question 1: Male and Female Graduate Enrollment (1967-2015)
```{r}
# Exploring the data

enroll_explore <- ggplot(enroll_tidy_df, aes(x = Year, y = Enrollment)) +
  geom_point(aes(color = Sex))

enroll_explore



# Female enrollment looks to be increasing at a much faster rate than male enrollment. Male enrollment was initially higher, but in the late 1980's, female enrollment surpassed male enrollement. Both enrollments level out at 2010. Male enrollment goes through stagnant periods and short spurts of increasing.


# Running a linear regression on dependent variable (y = male enrollment) by independent variale (x = year)

enroll_male_lm1 <- lm(`total Males` ~ Year, data = grad_enroll_df)

summary(enroll_male_lm1)

plot(enroll_male_lm1)

# Running a linear regression on dependent variable (y = female enrollment) by independent variable (x = year)

enroll_female_lm1 <- lm(`total Females` ~ Year, data = grad_enroll_df)

summary(enroll_female_lm1)

plot(enroll_female_lm1)

# To show the trends graphically, we will plot a graph over time of female and male enrollment to show the differences in trends.

grad_enroll_graph <- ggplot(enroll_tidy_df, aes(x = Year, y = Enrollment)) +
  geom_line(aes(color = Sex), size = 1) +
  theme_linedraw() +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0), breaks = seq(0, 2000000, 250000), limits = c(0,2000000)) +
  labs(x = "Year", y = "Total Graduate Enrollment", title ="Graduate Enrollment in the United States (1967-2015)") + scale_color_brewer(palette = "PuBuGn")

grad_enroll_graph
```


Question 2: Shifts in Female PhD Recipients by Field (1985,2000,2015)
```{r}

#Describe if and how there was a shift in PhDs awarded to females in four fields (Physical and Earth Sciences, Engineering, Education, and Humanities & Arts) in 1985, 2000, and 2015. Describe your results statistically, in a graph or table, and in text. 
#Tried to make a dataframe to graph the trends.. this was attempt 1:
phd_female_fields <- phd_field_df %>% 
  select("Field of study and sex", "Gender", "Number 1985", "Percent 1985", "Number 2000", "Percent 2000", "Number 2015", "Percent 2015") %>% 
  filter(Gender == "Female", `Field of study and sex` == "Physical sciences and earth sciences" | `Field of study and sex` == "Engineering" | `Field of study and sex` == "Education" | `Field of study and sex` == "Humanities and arts")
#but realized it was in a terrible format that R couldn't deal with so went back into excel and made a new one
#Attempt 2:

#reassign year to be a character

female_phd_tidy$Year <- as.character(female_phd_tidy$Year)

phd_female_number_graph <- ggplot(female_phd_tidy, aes(x = Year, y = `Number Enrolled`, group = `Field of study and sex`)) +
  geom_col(position = "dodge", aes(fill = `Field of study and sex`)) + 
  theme_classic() + 
  scale_y_continuous(expand = c(0,0)) + 
  scale_fill_discrete(name = "Field of Study") +
  labs(title = "Number of Female PhD Enrollments by Field (1985-2015)")
  
  
phd_female_number_graph

#Does this make sense, or is it better to group by year??? 
#I can't get the colors to change :(
#ok making another one that's based on percents  

phd_female_percent_graph <- ggplot(female_phd_tidy, aes(x = Year, y = `Percent Enrolled`, group = `Field of study and sex`)) +
  geom_col(position = "dodge", aes(fill = `Field of study and sex`)) +
   theme_linedraw() + 
  scale_fill_discrete(name = "Field of Study") +
  labs(title = "Percent of Female PhD Enrollments by Field (1985-2015)") +
  scale_fill_brewer(palette = "PuBuGn")
  
  
phd_female_percent_graph

#proptables and chi square


row.names(female_numbers_tidy) <- female_numbers_tidy$`Field of study and sex`
 
female_prop_final <- female_numbers_tidy %>% 
  select(`1985`, `2000`, `2015`) 


phd_chi <- chisq.test(female_prop_final)
phd_chi

#Ok proportions are significantly different, p < 2.2e-16

```
#Ran phd_chi$stdres in the console to find the significant differences. ALl are significantly different from expected. 
                                           1985       2000       2015
Physical sciences and earth sciences -10.376557  -8.559263  17.031216
Engineering                          -24.746239 -12.702359  33.184446
Education                             29.337381   7.576085 -32.127993
Humanities and arts                   -5.666481   7.947822  -2.866323



Question 3: Male and Female Salaries for Starting PostDoc & Other Employment Positions
```{r}
phd_med <- phd_med_sal_df %>% 
  filter(Employment != "Male")
View(phd_med)

phd_med$Employment <- factor(phd_med$Employment)
phd_med$X3 <- factor(phd_med$X3)
phd_med$`Postdoctoral study` <- factor(phd_med$`Postdoctoral study`)
phd_med$X5 <- factor(phd_med$X5)

# Running paired t-tests for the median salaries of male and females

# Regular employment

# It doesn't work! need to do mann whitney U/wilcox


phd_med_sal_tidy_df

employment_t <- t.test()
```
```{r}
# Question 3

# Exploratory Analysis

# Make histogram of median salary vs sex, Employment
employment_test_df <- phd_med_sal_tidy_df %>% 
  filter(Planas == "EMPLOYMENT")

employment_test_hist <- ggplot(employment_test_df, aes(x = Median.Salary )) +
  geom_histogram(bins = 6, aes(fill = Sex)) +
  facet_wrap(~ Sex, scale = "free") +
  theme_classic() +
  theme(legend.position = "none") + 
  scale_y_continuous(expand = c(0,0))

employment_test_hist

# Make histogram of median salary vs sex, Employment
postdoc_test_df <- phd_med_sal_tidy_df %>% 
  filter(Planas == "POSTDOC")

postdoc_test_hist <- ggplot(postdoc_test_df, aes(x = Median.Salary )) +
  geom_histogram(bins = 6, aes(fill = Sex)) +
  facet_wrap(~ Sex, scale = "free") +
  theme_classic() +
  theme(legend.position = "none") + 
  scale_y_continuous(expand = c(0,0))

postdoc_test_hist


employment_mannu <- wilcox.test(phd_med$Employment, phd_med$X3, paired = TRUE, data = phd_med)

# didn't work

med_emp_male <- c(78000, 75000, 75000, 80000, 75167, 95000, 105000, 63000, 105000, 64000, 95000, 71000, 52000, 123500, 62800)

med_emp_fem <- c(66000, 66000, 75000, 75000, 71750, 97650, 90000, 60000, 95750, 62000, 90000, 63000, 50000, 120000, 61000)

employment_wilcox <- wilcox.test(med_emp_male, med_emp_fem, exact = FALSE, paired = TRUE)

employment_wilcox

med_post_male <- c(42750, 42000, 43000, 42000, 50000, 50000, 58000, 42000, 65000, 48000, 45000, 50000, 45000, 60000, 50000)

med_post_fem <- c(44000, 42000, 43250, 42000, 50000, 53000, 55000, 42000, 65000, 49250, 45000, 45000, 45000, 63500, 44000)

postdoc_wilcox <- wilcox.test(med_post_male, med_post_fem, exact = FALSE, paired = TRUE)

postdoc_wilcox

# Column Graphs for Visualization

phd_med_sal_tidy_df2 <- phd_med_sal_tidy_df %>% 
  mutate(sex = ifelse(Sex == "FEMALE", "Male", "Felmale")) %>% 
  select(Field, Median.Salary, Planas, sex)

phd_med_tidy_employ <- phd_med_sal_tidy_df2 %>% 
  filter(Planas == "EMPLOYMENT")

employ_graph <- ggplot(phd_med_tidy_employ, aes(x = Field, y = Median.Salary)) +
  geom_col(position = "dodge", aes(fill = sex)) +
   theme_linedraw() + 
  scale_fill_discrete(name = "Sex") +
  coord_flip() +
   scale_y_continuous(expand = c(0,0), breaks = seq(0, 125000, 25000), limits = c(0,125000)) +
  labs(title = "Median Starting Salaries for PhD Recipients in non-PostDoc Positions", x = "Field", y = "Median Salary ($)") 

employ_graph


phd_med_tidy_post <- phd_med_sal_tidy_df2 %>% 
  filter(Planas == "POSTDOC")

postdoc_graph <- ggplot(phd_med_tidy_post, aes(x = Field, y = Median.Salary)) +
  geom_col(position = "dodge", aes(fill = sex)) +
   theme_linedraw() + 
  scale_fill_discrete(name = "Sex") +
  coord_flip() + 
  scale_y_continuous(expand = c(0,0), breaks = seq(0, 80000, 20000), limits = c(0,80000)) +
  labs(title = "Median Starting Salaries for PhD Recipients in PostDoc Positions", x = "Field", y = "Median Salary ($)")

postdoc_graph
```


Question 4: Exploring Academic Salaries for Professors in U.S. Colleges
```{r}

# Create Histogram Years Since PhD vs Salary

years_since_phd_gph <- ggplot(fac_sal_df, aes(x = Years.Since.PhD, y = Salary)) +
  geom_point(aes(color = Sex), alpha = 0.5) +
  facet_wrap(~ Faculty.Rank)

years_since_phd_gph
```




```{r}
#Linear regession exploration 
pairs_df <- fac_sal_df %>% 
  select(Years.Since.PhD, Years.Faculty.Service, Salary)

pairs(pairs_df)
#Salary ad years since PHD looks more linear thean Years.faculty.service

salary_lm <- lm(Salary ~ Sex + Years.Faculty.Service + Years.Since.PhD + Discipline + Faculty.Rank, data = fac_sal_df)
salary_lm

# b. Model summary:

summary(salary_lm)
AIC(salary_lm)
# AIC = 9093.826

#Visualize residuals
plot(salary_lm)

vif(salary_lm)
#ok yikes so years of faculty service and years since PhD are super correlated. 
#I feel like basically we only use one of those I feel like that's fine since most people will work right after their phd, doesn't make sense to include both. 


salary_lm2 <- lm(Salary ~ Sex + Years.Since.PhD + Discipline + Faculty.Rank, data = fac_sal_df)
salary_lm2

summary(salary_lm2)

AIC(salary_lm2)
#AIC = 9097.22

plot(salary_lm2)
#resdiuals are bad, but the qq plot looks good. 

vif(salary_lm2)
#woooo all are under 4 so we are good to go re: colinearlity 


#my prediction is that male/female doesn't change the slope of the line, they just have differne t starting point, so we need that interation term to make up for that? 
#I will graph it to see if it's visual
#ok no that predcition was wrong. Sex doesn't appear to be that significant, there's just way more males than females, and way more old males. 

#For exploratory purposes, what about not including sex? 
salary_lm3 <- lm(Salary ~ Years.Since.PhD + Discipline + Faculty.Rank, data = fac_sal_df)
salary_lm3
AIC(salary_lm3)
#AIC = 9096.497
plot(salary_lm3)
vif(salary_lm3)

#For exploratory purposes, maybe remove faculty rank, since that goes along with years since PHD?

salary_lm4 <- lm(Salary ~ Years.Since.PhD + Discipline, data = fac_sal_df)
salary_lm4
AIC(salary_lm4)
#AIC = 9217.566
plot(salary_lm4)
vif(salary_lm4)


gender_dif_graph <- ggplot(fac_sal_df, aes(x = Years.Since.PhD, y = Salary, group = Sex)) + 
  geom_point(aes(color = Faculty.Rank, shape = Discipline)) +
  facet_wrap(~Sex)

gender_dif_graph
                           



```

